---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoptions,echo=TRUE}
knitr::opts_chunk$set(echo=TRUE,warning = F,message=F)
```


## Loading and preprocessing the data

```{r}
library(readr)
options(digits = 2)
activity=read_csv("activity.csv",col_names = T)
str(activity)
```


## What is mean total number of steps taken per day?
```{r}
library(dplyr)
activitytotal=activity %>% 
    group_by(date) %>% 
    summarize(steps_total=sum(steps,na.rm = T))
mean=mean(activitytotal$steps_total)
median=median(activitytotal$steps_total)
library(ggplot2)
ggplot(activitytotal,aes(date,steps_total))+
    geom_histogram(stat="identity")+
    labs(x="Date",y="Total steps",title="Total number of steps taken per day")

```

The mean total number of steps taken per day was `r mean`, and the median number was `r median`.

## What is the average daily activity pattern?
```{r pattern}
activitymean=activity %>% 
    group_by(interval) %>% 
    summarize(meansteps=mean(steps,na.rm=T))
maxsteps=max(activitymean$meansteps)
maxinterval=activitymean$interval[activitymean$meansteps==maxsteps]
ggplot(activitymean,aes(interval,meansteps))+
    geom_line(col="blue")+
    labs(x="Interval",y="Mean steps",title="The average number of steps taken per interval")
```

The `r maxinterval` interval, on average across all the days, contains the maximum number of steps, which was `r maxsteps` steps.

## Imputing missing values
```{r}
na=sum(is.na(activity$steps))
activityimpute = activity %>% 
    group_by(interval) %>% 
    mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=T),steps))
activityimputetotal=activityimpute %>% 
    group_by(date) %>% 
    summarize(steps_total=sum(steps))
mean_i=mean(activityimputetotal$steps_total)
median_i=median(activityimputetotal$steps_total)
ggplot(activityimputetotal,aes(date,steps_total))+
    geom_histogram(stat="identity")+
    labs(x="Date",y="Total steps",title="Total number of steps taken per day based on imputed data")
```

Based on imputed data, the mean total number of steps taken per day was `r mean_i`, and the median number was `r median_i`. The mean and median of imputed data are a little bit higher than data with NA.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lubridate)
activityimpute=activityimpute %>% 
    mutate(weekend=factor(ifelse(wday(date)==1|wday(date)==7,1,0),levels=c(1,0),labels = c("Weekend","weekday")))
activityimputemean=activityimpute %>% 
    group_by(weekend,interval) %>% 
    summarise(meansteps=mean(steps))
ggplot(activityimputemean,aes(interval,meansteps))+
    geom_line(col="blue")+
    facet_wrap(~weekend)+
    labs(x="Interval",y="Mean steps",title="The average number of steps taken per interval on weekdays and weekends")
```


